### 2015-07-06 narrative/qos.tex
one pass on qos.

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
If the system uses reservations intelligently, its performance is predictable and
efficient because resources are not wasted on interference. However,
in cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and reverting back to the previous committed resource level. This can
be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

@@

Performance isolation (including its utilization and charging models)
critically depends on \emph{workload-independent metrics}. Even
though throughput and latency are frequently the most meaningful
metrics for an application, they are not workload-independent and
therefore are inadequate for measuring the utilization of a device
across all workloads, unless one accepts worst-case estimates that
can lead to gross underutilization of resources (e.g. the throughput
of random and sequential workloads in spinning media differ by
orders of magnitudes). A common workload-independent metric is
\emph{time utilization}, i.e. the amount of time a resource is
utilized in a given time interval. Time utilization has the advantage
of having an always defined maximum and therefore makes resources
fully reservable. Even though a workload-independent metric might
at first not appear meaningful for an application, given a particular
workload an application can discover the relationship between its
workload-dependent metric and the workload-independent metric.
Because of performance isolation, the application has to discover
this relationship only once.

To achieve effective performance isolation, we seek to expand the resource management
capabilities in Sirocco to support additional metrics to guide resource management
decisions. Currently Sirocco offers automatic resource management capabilities,
based on resource capacity and data resilience requirements. 
 Our basic goal is to offer data-centric metrics that help guide the
forced resource management decisions.

@@

\paragraph{Related Work:} Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.

>>>>

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
If the system uses reservations intelligently, its performance is predictable and
efficient because resources are not wasted on interference.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest. This can
be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

@@

Performance isolation (including its utilization and charging models)
critically depends on \emph{workload-independent metrics}. Even
though throughput and latency are frequently the most meaningful
metrics for an application, they are not workload-independent and
therefore are inadequate for measuring the utilization of a device
across all workloads, unless one accepts worst-case estimates that
can lead to gross underutilization of resources (e.g. the throughput
of random and sequential workloads in spinning media differ by
orders of magnitudes). A common workload-independent metric is
\emph{time utilization}, i.e. the amount of time a resource is
utilized in a given time interval. Time utilization has the advantage
of having an always defined maximum and therefore makes resources
fully reservable. Even though a workload-independent metric might
at first not appear meaningful for an application, given a particular
workload an application can discover the relationship between its
workload-dependent metric and the workload-independent metric.
Because of performance isolation, the application has to discover
this relationship only once.
To achieve effective performance isolation, we seek to expand the resource management
capabilities in Sirocco to support additional utilization 
metrics to guide resource management
decisions. Currently, Sirocco offers automatic resource management capabilities,
based on resource capacity and data resilience requirements, and it needs to be extended
to offer data-centric metrics that help guide the forced QoS decisions.

@@

\paragraph{Related Work:} On current HPC systems, absolute IO performance
guarantees are not possible. Previous work \cite{lofstead:2010:adaptive,liu_hotstorage} 
from this project team uses a client-assisted approach to mitigate IO variablity.
Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.
